\Title

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
task

Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays. In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay as a computational process. As a consequence, this extends our previous model by encoding spiking patterns as 3D images. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays.
%
\subsection{Task definition: fast motion detection}
%
To test our model, we will quantify its ability to categorize different motions. In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\cite{leon2012motion} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}). In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$. Movies' duration are here set to $\Ntime=400$.
This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions:
% of a motion cloud is determined by a constant $k_v$: with a constant speed $\speed$ (see \fig{motionclouds}-(a)): 
%\begin{equation*}
%%\begin{gathered}
%\va{v}_k = %\binom{\cos(\frac{2\pi\cdot k_v}{N_v})}{\sin(\frac{2\pi\cdot k_v}{N_v})}
%  \begin{pmatrix}
%    \cos(2\pi\cdot \frac{k}{\Nclass})\\
%    \sin(2\pi\cdot \frac{k}{\Nclass})
%  \end{pmatrix}
%%\end{gathered}
%\end{equation*}
$
\speed_\kernelind = 
  ( 
    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
  )
$ (\seeFig{motionclouds}-(a) for an illustration).
For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content. This method provides a rich dataset of textured movies for which we know the ground truth for motion.
%
\begin{figure}[h!]%[ht!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
    \caption{{\bf Motion detection task.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red. {\bf (b)} A screenshot of one generated naturalistic textured stimulus at a specific time. % $\timev_\arank$. 
    {\bf (c)} The corresponding ON (in red) and OFF (in blue) event stream generated from the stimuli on (b) and constituting the input to the spiking neural network.}
    \label{fig:motionclouds}
\end{figure}

@@

%
%\begin{figure}[t!]
%    \centering
%    \includegraphics[width=0.90\linewidth]{figures/figure_1a.pdf}
%    \includegraphics[width=0.90\linewidth]{figures/figure_1b.pdf}
%    \includegraphics[width=0.90\linewidth]{figures/figure_1c.pdf}
%	    \caption{The hetero-synaptic delays model applied on synthetic data. {\bf (a)}~The afferent information consists of the repeated occurrence of groups of precise motifs of spikes that we call ``polychronous groups'' (PGs). We highlight them by different colors, an information hidden to a detection model. {\bf (b)}~The model is defined as an assembly of neurons (here for $4$ PGs) each defined by a set of different synapses defined by weights (increasing with the radius of the black dots) at each different delay. Note that each afferent may be connected with multiple weights at different delays. The propagation of the afferent information through these delay may generate at each time step a synchronous pattern on a subset of synapses. %These PG patterns were efficiently learned by a logistic regression.
%	    {\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG pattern at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences. 
%	    \noteAG{clarify the figure - size of the dots for the synaptic weights - how the delay is related to the time we observe the spike trains?}
%	    }
%	    \noteLP{cette figure me parait superflue maintenant, et elle n'eest pas citée dan sle corps du texte - on pourrait simplement l'enlever. c'est la seule aussi (avec le paragraphe en entrée) qui parle de "polychronous groups" - l'enlever permet de simplifier l'ensemble du texte. }
%    \label{fig:model}
%\end{figure}
%
\subsection{Hetero-synaptic delays model}
%
\begin{figure*}[ht!]
    {\centering
    \vspace{-3cm}
    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
    }
    %\vspace{-.5cm}
    \caption{
    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities). These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown. One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
	}
    \label{fig:kernels}
\end{figure*} 
%
The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$. 
This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).

@@

In Figure~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer. As a comparison and to account for the gain in performance by using hetero-synaptic delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\cite{Grimaldi22pami}. This latter method is based on delays from the last recorded events and uses fewer computations (in our case $15\times15$) than the dense 3D kernels without any pruning ($15\times15\times8$). While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights follows a sigmoid curve. Half-saturation level is reached at about $3.5\times 10^{-3}\%$ of active weights, corresponding to a total amount of $6$ computations per decision. Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $200$. In this case, the number of computations is greatly reduced compared to~\cite{Grimaldi22pami}, thus demonstrating the efficiency of the presented method. 
%
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/accuracy.pdf}
    \caption{Accuracy as a function of the number of computation load for the hetero-synaptic delays model (in blue) and for a method using 2D time surfaces (in red). The relative computational load (on a log axis) is controlled by changing the percentage of active weights relative to the dense convolution kernel. We observe a similar accuracy than HOTS, yet that our model could achieve a similar accuracy with significantly fewer coefficients.}
    \label{fig:accuracy}
\end{figure}
%
\section{Discussion}

>>>>

Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays. In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay as a computational process. As a consequence, this extends our previous model by encoding spiking patterns as 3D images. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays. %
%
We expect our model to be able to efficiently detect temporal patterns and here we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
%
\subsection{Task definition: fast motion detection}
%
Taking analogy with biological conditions, a common use case for motion detection is to estimate the motion generated on a sensor by saccade-like movements. Saccades are rapid movements of the eyeball which allow to reposition the axis of view in a short period. In humans, these are particularly rapid 
To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural

use a set of synthetic visual stimuli, \textit{Motion Clouds}~\cite{leon2012motion} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}). In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$. Movies' duration are here set to $\Ntime=400$.
This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions:
% of a motion cloud is determined by a constant $k_v$: with a constant speed $\speed$ (see \fig{motionclouds}-(a)): 
%\begin{equation*}
%%\begin{gathered}
%\va{v}_k = %\binom{\cos(\frac{2\pi\cdot k_v}{N_v})}{\sin(\frac{2\pi\cdot k_v}{N_v})}
%  \begin{pmatrix}
%    \cos(2\pi\cdot \frac{k}{\Nclass})\\
%    \sin(2\pi\cdot \frac{k}{\Nclass})
%  \end{pmatrix}
%%\end{gathered}
%\end{equation*}
$
\speed_\kernelind = 
  ( 
    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
  )
$ (\seeFig{motionclouds}-(a) for an illustration).
For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content. This method provides a rich dataset of textured movies for which we know the ground truth for motion.
%
\begin{figure}[h!]%[ht!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
    \caption{{\bf Motion detection task.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red. {\bf (b)} A screenshot of one generated naturalistic textured stimulus at a specific time. % $\timev_\arank$. 
    {\bf (c)} The corresponding ON (in red) and OFF (in blue) event stream generated from the stimuli on (b) and constituting the input to the spiking neural network.}
    \label{fig:motionclouds}
\end{figure}

@@

\subsection{Hetero-synaptic delays model}
%
\begin{figure*}[ht!]
    {\centering
    \vspace{-3cm}
    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
    }
    %\vspace{-.5cm}
    \caption{
    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities). These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown. One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
	}
    \label{fig:kernels}
\end{figure*} 
%
The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$. 
This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).

@@

In Figure~\ref{fig:accuracy}, we plot the classification accuracy as a function of the relative number of computations, or active weights, per decision for each neuron of the layer. As a comparison and to account for the gain in performance by using hetero-synaptic delays, we provide the accuracy obtained with a MLR model using 2D time surface (in red) as in~\cite{Grimaldi22pami}. This latter method is based on delays from the last recorded events and uses fewer computations (in our case $15\times15$) than the dense 3D kernels without any pruning ($15\times15\times8$). While less computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.
By pruning weights, we observe that the evolution of accuracy as a function of the log percentage of active weights follows a sigmoid curve. Half-saturation level is reached at about $3.5\times 10^{-3}\%$ of active weights, corresponding to a total amount of $6$ computations per decision. Compared to the full kernels, the accuracy of our method is maintained to its top performances when dividing the number of computations by a factor up to about $200$. In this case, the number of computations is greatly reduced compared to~\cite{Grimaldi22pami}, thus demonstrating the efficiency of the presented method. 
%
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/accuracy.pdf}
    \caption{Accuracy as a function of the number of computation load for the hetero-synaptic delays model (in blue) and for a method using 2D time surfaces (in red). The relative computational load (on a log axis) is controlled by changing the percentage of active weights relative to the dense convolution kernel. We observe a similar accuracy than HOTS, yet that our model could achieve a similar accuracy with significantly fewer coefficients.}
    \label{fig:accuracy}
\end{figure}
%

\subsection{Testing with natural-like textures}
%
To test our model, we will quantify its ability to categorize different motions. In that order, we use a set of synthetic visual stimuli, \textit{Motion Clouds}~\cite{leon2012motion} which are natural-like random textures for which we can control for velocity, among other parameters (\seeFig{motionclouds}). In particular, we will set the spatial size to $(\Nx, \Ny)$ and consider a discretization of time with a time step of $1$ such that $\timev \in \mathbb{N}$. Movies' duration are here set to $\Ntime=400$.
This procedure defines a set of textures with different spatial properties and different motions $\va{v}_k$ with  $1 \le k \le \Nclass$ and $\Nclass=8$ defined by a constant speed and linearly spaced directions:
% of a motion cloud is determined by a constant $k_v$: with a constant speed $\speed$ (see \fig{motionclouds}-(a)): 
%\begin{equation*}
%%\begin{gathered}
%\va{v}_k = %\binom{\cos(\frac{2\pi\cdot k_v}{N_v})}{\sin(\frac{2\pi\cdot k_v}{N_v})}
%  \begin{pmatrix}
%    \cos(2\pi\cdot \frac{k}{\Nclass})\\
%    \sin(2\pi\cdot \frac{k}{\Nclass})
%  \end{pmatrix}
%%\end{gathered}
%\end{equation*}
$
\speed_\kernelind = 
  ( 
    \speed \cdot \cos(2\pi\cdot \frac{\kernelind}{\Nspeed}),
    \speed \cdot \sin(2\pi\cdot \frac{\kernelind}{\Nspeed})
  )
$ (\seeFig{motionclouds}-(a) for an illustration).
For any given velocity, we also varied the parameters of the textures, such as the mean and variance of the orientation or spatial frequency content. This method provides a rich dataset of textured movies for which we know the ground truth for motion.
%
\begin{figure}[h!]%[ht!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/motionclouds.pdf}
    \caption{{\bf Motion detection task.} {\bf (a)} The motion direction represented as the plain red vector, other possible motion directions are represented in light red. {\bf (b)} A screenshot of one generated naturalistic textured stimulus at a specific time. % $\timev_\arank$. 
    {\bf (c)} The corresponding ON (in red) and OFF (in blue) event stream generated from the stimuli on (b) and constituting the input to the spiking neural network.}
    \label{fig:motionclouds}
\end{figure}

\section{Discussion}
